Describe and Explore: Using feature-based time series clustering to identify meaningful structures in intensive longitudinal data

### 2023-01-17 methods-and-results.Rmd
:shit: formula tst

For the central tendency: $Median$ where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$.

\begin{equation} \label{eq:median}
  Median(X_{ij}) = 
    \begin{cases}
      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
    \end{cases}
\end{equation}

For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the $Median$ (as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$.

\begin{equation} \label{eq:mad}
  MAD(x_{ijt}) = Median(| x_{ijt} - \overline{x_{ij}} |)
\end{equation}

>>>>

For the central tendency: \textit{median} ($M$) where $X_{ij}$ is the an ordered list of values from the time series of variable $j$ for participant $i$. The calculation depends on whether the number of measurements in a time series $n$ is odd or even.

\begin{equation} \label{eq:median}
  M(X_{ij}) = 
    \begin{cases}
      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
    \end{cases}
\end{equation}

For distribution we chose the \textit{Median Absolute Deviation} ($MAD$), where we calculate the \textit{median} ($M$; calculated as in \eqref{eq:median}) for the absolute deviations of measurement $x$ at time point $t$ for participant $i$ and variable $j$ from the median of that time series $X$.

\begin{equation} \label{eq:mad}
  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
\end{equation}

mean absolute change (MAC) for stability

\begin{equation} \label{eq:mac}
  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
\end{equation}

lag--1 autocorrelation for inertia: lag $l=1$

\begin{equation} \label{eq:ar}
  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
\end{equation}

overall linear regression slope to describe the linear trend

summarized the nonlinear trend with the estimated degrees of freedom of an empty GAM spline model (edf) --- the edf summarizes the \textit{wiggliness} of the spline trend line.

### 2023-01-17 methods-and-results.tex
:shit: formula tst

For the central tendency: \(Median\) where \(X_{ij}\) is the an ordered
list of values from the time series of variable \(j\) for participant
\(i\).

\begin{equation}
  Median(X_ij) = \left\{ 
    \begin{array}{ll}
      X \left[ \frac{n+1}{2} \right] & if n is odd \\
      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & if n is even
    \end{array}
\end{equation}

>>>>

For the central tendency: \textit{median} (\(M\)) where \(X_{ij}\) is
the an ordered list of values from the time series of variable \(j\) for
participant \(i\). The calculation depends on whether the number of
measurements in a time series \(n\) is odd or even.

\begin{equation} \label{eq:median}
  M(X_{ij}) = 
    \begin{cases}
      X \left[ \frac{n+1}{2} \right] & \text{if $n$ is odd} \\
      \frac{X \left[ \frac{n}{2} \right] + X \left[ \frac{n}{2} +1 \right]}{2} & \text{if $n$ is even}
    \end{cases}
\end{equation}

For distribution we chose the \textit{Median Absolute Deviation}
(\(MAD\)), where we calculate the \textit{median} (\(M\); calculated as
in \eqref{eq:median}) for the absolute deviations of measurement \(x\)
at time point \(t\) for participant \(i\) and variable \(j\) from the
median of that time series \(X\).

\begin{equation} \label{eq:mad}
  MAD(X_{ij}) = M(\left| x_{ijt} - M(X_{ij})} \right|)
\end{equation}

mean absolute change (MAC) for stability

\begin{equation} \label{eq:mac}
  MAC(X_{ij}) = \frac{1}{n-1} \sum_{t=1, \ldots, t-1}\left|x_{t+1}-x_t\right|
\end{equation}

lag--1 autocorrelation for inertia: lag \(l=1\)

\begin{equation} \label{eq:ar}
  \frac{1}{(n-l)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+l}-\mu) \\
  \frac{1}{(n-1)\sigma^2}\sum_{t=1}^{n-1}(X_t-\mu)(X_{t+1}-\mu)
\end{equation}

overall linear regression slope to describe the linear trend

summarized the nonlinear trend with the estimated degrees of freedom of
an empty GAM spline model (edf) --- the edf summarizes the
\textit{wiggliness} of the spline trend line.
